Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-22 main.tex
fonet -> MLS

A rectified derivation $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ determines a
substitution which maps the existential bound variables occurring in
$A$ to the terms substituted for them in the instances of the
$\exists$-rule in $\Phi$. We denote this substitution with
$\dsubstof\Phi$ and call it the \bfit{dualizer} of $\Deri$. Furthermore, every atom occuring in the conclusion $A$
must be consumed by a unique instance of the rule $\aiD$ in
$\Deri$. This allows us to define a (partial) equivalence relation
$\linkingof\Deri$ on the atom occurrences in $A$ by $a\linkingof\Deri
b$ if $a$ and $b$ are consumed by the same instance of $\aiD$
in~$\Deri$. We call $\linkingof\Deri$ the \bfit{linking} of~$\Deri$.

@@

\begin{thm}\label{thm:MLS1->fonet}
  Let $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ be a recified derivation. Then
  $\tuple{\graphof{A},\linkingof{\Deri}}$ is fonet and $\dsubstof{\Deri}$ is a dualizer for it.
\end{thm}

@@

\begin{thm}
  \label{thm:fonet->MLL}
  Let $\tuple{\gC,\linkingof\gC}$ be a fonet, and let $\Gamma$ be a sequent with $\graphof\Gamma=\gC$. Then there is an
  $\FOMLL$-proof $\Pi$ of $\Gamma$, such that
  $\fographof\Pi=\tuple{\gC,\linkingof\gC}$.
\end{thm}

@@

\begin{thm}
  Let $\gC$ be a fonet with linking $\linkingof\gC$, let $\dualizer$ be a dualizer
  for $\gC$, and let $A$ be a formula with $\graphof A=\gC$.  Then
  there is a derivation $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ such that
  $\linkingof\gC=\linkingof{\Deri}$ and $\dualizer=\substof{\Deri}$.
\end{thm}

\begin{proof}
  \todo{}
\end{proof}

>>>>

A rectified derivation $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ determines a
substitution which maps the existential bound variables occurring in
$A$ to the terms substituted for them in the instances of the
$\exists$-rule in $\Phi$. We denote this substitution with
$\dsubstof\Phi$ and call it the \bfit{dualizer} of
$\Deri$. Furthermore, every atom occuring in the conclusion $A$ must
be consumed by a unique instance of the rule $\aiD$ in $\Deri$. This
allows us to define a (partial) equivalence relation $\linkingof\Deri$
on the atom occurrences in $A$ by $a\linkingof\Deri b$ if $a$ and $b$
are consumed by the same instance of $\aiD$ in~$\Deri$. We call
$\linkingof\Deri$ the \bfit{linking} of~$\Deri$, and define
$\fographof\Deri=\tuple{\graphof{A},\linkingof{\Deri}}$.

@@

\begin{thm}\label{thm:MLS1->fonet}
  Let $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ be a recified derivation. Then
  $\fographof\Deri$ is a fonet and $\dsubstof{\Deri}$ a dualizer for it.
\end{thm}

@@

\begin{thm}
  \label{thm:fonet->MLL1}
  Let $\tuple{\gC,\linkingof\gC}$ be a fonet, and let $\Gamma$ be a sequent with $\graphof\Gamma=\gC$. Then there is an
  $\FOMLL$-proof $\Pi$ of $\Gamma$, such that
  $\fographof\Pi=\tuple{\gC,\linkingof\gC}$.
\end{thm}

@@

We can now straightforwardly obtain the same result for $\FOMLS$:

\begin{thm}
  \label{thm:fonet->MLS1}
  Let $\tuple{\gC,\linkingof\gC}$ be a fonet, and let $C$ be a formula with $\graphof C=\gC$. Then there is
  a derivation $\vlder{\FOMLS}{\Deri}{C}{\ttt}$ such that $\fographof\Deri=\tuple{\gC,\linkingof\gC}$.
  %% an
  %% $\FOMLL$-proof $\Pi$ of $\Gamma$, such that
  %% $\fographof\Pi=\tuple{\gC,\linkingof\gC}$.
  %% Let $\gC$ be a fonet with linking $\linkingof\gC$, let $\dualizer$ be a dualizer
  %% for $\gC$, and let $A$ be a formula with $\graphof A=\gC$.  Then
  %% there is a derivation $\vlder{\FOMLS}{\Deri}{A}{\ttt}$ such that
  %% $\linkingof\gC=\linkingof{\Deri}$ and $\dualizer=\substof{\Deri}$.
\end{thm}

\begin{proof}
  We apply Theorem~\ref{thm:fonet->MLL1} to obtain a sequent
  proof~$\Pi$ of $\sqn C$ with
  $\fographof\Pi=\tuple{\gC,\linkingof\gC}$. Then we apply
  Lemma~\ref{lem:MLL1->MLS1}, observing that the translation from
  $\FOMLL$ to $\FOMLS$ preserves linking and dualizer.
\end{proof}

\begin{remark}
  Note that it is also possible to do a direct ``sequentialization''
  into the deep inference system $\FOMLS$, using the techniques
  presented in \cite{dissvonlutz} and \cite{str:MLL2}.
\end{remark}
